Clique and Constraint Models for Maximum Common (Connected) Subgraph Problems

### 2016-04-14 paper/paper.tex
Move figure back to first page
Bring in Patrick's text

\begin{figure}[t] % first page if possible
    \centering
    \begin{tikzpicture}[scale=0.33]%{{{
        \begin{scope}[rotate=90]
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!40, fill=white, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};

@@

\paragraph{Ensuring connectedness with a specific filtering.}
Another way to ensure connectedness is to filter domains to ensure connectedness consistency. Let us consider two graphs $G$ and $G'$, and let $D$ be the current domains (we suppose that $D(x_u)$ is a singleton when $x_u$ is assigned). Let $C$ (resp. $C'$) be the set of nodes of $G$ (resp. $G'$) which may belong to the common subgraph, {\em i.e.}, $C=\{ u\in N_{G}, D(x_u)\neq\{\bot\}\}$ (resp. $C' = \{u'\in N_{G'}, \exists u\in N_G, u'\in D(x_u)\}$). Let $G_{\downarrow C}$ (resp. $G'_{\downarrow C'}$) be the subgraph of $G$ induced by $C$ (resp. the subgraph of $G'$ induced by $C'$).
The connectedness consistency ensures that both $G_{\downarrow C}$ and $G'_{\downarrow C'}$ are connected graphs.

@@

\christine{We have to relate this basic filtering with existing work: Connectedness \cite{Brown:2005} Reachability \cite{DBLP:conf/cp/DoomsDD05,DBLP:conf/cp/QuesadaRD05}.

Samba and I have tried to find more clever ways for doing this filtering (in an incremental way). But we failed in finding something cheaper than performing a traversal each time a set of values are removed. In particular, we tried to adapt ideas from the paper of Patrick, but the problem is that we have to maintain counters on edges, and these counters must be restored when backtracking, and this is as expensive as performing a traversal (?). Any brilliant idea to share with us on this point?}

>>>>

\begin{figure}[b] % first page if possible
    \centering
    \begin{tikzpicture}[scale=0.33]%{{{
        \begin{scope}[rotate=90]
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!40, fill=white, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};

@@

\paragraph{Ensuring connectedness with a specific filtering.}

CP(Graph) \cite{DBLP:conf/cp/DoomsDD05} introduces graph domain variables and enforces connectivity via the
\emph{reachable} constraint, ensuring that there is a path from a specified vertex to a specified
set of vertices. One such constraint is then posted for each of the vertices in the graph and
encodes the transitive closure of the graph. Brown et al.\ \cite{Brown:2005} explored the use of
constraint programming in the generation of connected graphs with specified degree sequences.  Two
constraints were combined, i.e.\ the graphical constraint (a backtrackable implementation of the
Havel-Hakimi algorithm) with a connectivity constraint implemented using sets of vertices, where
vertex sets $A$ and $B$ are combined when there exists a pair of vertices $v \in A$ and $w \in B$
and an edge $(v,w) \in E$. Residual degree counts are maintained on components and vertices to
enforce graphicality and connectivity. In \cite{DBLP:conf/ecai/ProsserU06} a connectivity constraint was proposed for
connected graph generation where decision variables are edges (the search process accepts and
rejects edges). The constraint employed depth first search to maintain the set of tree edges and
back edges, associating path counters on these edges. The counters were then used to detect the
existence of cut-edges and protects these by forcing edges. The complexity of this constraint is
cubic in the number of vertices.

In all these previous works, the goal is to ensure that a given set of vertices is connected. For
MCCS, the problem is slightly different: we have to ensure that the number of connected vertices
that may be matched (in both graphs) is greater than the size of the largest common subgraph
previously found. Therefore, we introduce a new filtering algorithm to ensure connectedness
consistency for MCCS.  Let us consider two graphs $G$ and $G'$, and let $D$ be the current domains
(we suppose that $D(x_u)$ is a singleton when $x_u$ is assigned). Let $C$ (resp. $C'$) be the set of
nodes of $G$ (resp. $G'$) which may belong to the common subgraph, {\em i.e.}, $C=\{ u\in N_{G},
D(x_u)\neq\{\bot\}\}$ (resp. $C' = \{u'\in N_{G'}, \exists u\in N_G, u'\in D(x_u)\}$). Let
$G_{\downarrow C}$ (resp. $G'_{\downarrow C'}$) be the subgraph of $G$ induced by $C$ (resp. the
subgraph of $G'$ induced by $C'$).  The connectedness consistency ensures that both $G_{\downarrow
C}$ and $G'_{\downarrow C'}$ are connected graphs.

@@
